Push-Pull Signal-Function Functional Reactive Programming

### 2012-11-30 Docs/IFL2012/implementation.tex
ifl draft edits

The design of signal functions specifies a family of types for the inputs and
outputs of signal functions. Signal functions are not functions in the purest
sense, however. They are not mappings from a single instance of their input
type to a single instance of their output type. They must be implemented with
respect to the temporal semantics of their inputs and outputs.

We therefore start by creating a set of concrete datatypes for the inputs and
outputs of signal functions. These datatypes will be parameterized by the input
and output types of the signal function, and will not be exposed to the user of
the library. Rather, they will specify how data is represented during the
temporal evaluation of signal functions. We then implement signal functions
as records of functions from these concrete types to these concrete types paired
with new signal functions. The evaluation system for signal functions then
maintains this record for the signal function it is evaluating, calls the
correct function for the current input, actuates the outputs (using supplied
functions from the outputs to monadic actions) and replaces the record with the
newly supplied signal function record.

In Section~\ref{section:System_Design_and_Interface} we presented signal vectors
as a set of types. In order to be completely principled, we should isolate these
types into their own {\em kind} (a sort of type of types); however, the Haskell
extension for this was far from stable at the time this system was created.

Using the Glasgow Haskell Compiler's extension for Generalized Algebraic
Datatypes~\cite{Cheney2003,Xi2003,PeytonJones2006}, we specify three concrete
datatypes which are parameterized over signal vectors, and represent
information about one temporal point in a signal function's evaluation.

@@

The next representation contains a constructor for {\tt SVEvent} nodes, carrying
a value typed by the type parameter of {\tt SVEvent}, and two constructors, each
of which carries a representation for one child of a {\tt SVAppend} node (the
left or right), and is parametric in the other. Thus, an instance of this
representation is a path from the root of a signal vector to an event node,
carrying exactly one value for the event.

>>>>

The design of signal functions specifies a family of types for the inputs and
outputs of signal functions. ``Signal functions'' are not simple functions,
rather, they are transformers of signals and events. They are reactive, that is,
they can alter themselves in response to their inputs.

We start by creating a set of concrete datatypes for the inputs and outputs of
signal functions. These datatypes will be parameterized by the input and output
signal vectors of the signal function, and will not be exposed to the user of
the library. Rather, they will specify how data is represented during the
evaluation of signal functions. We then implement signal functions
as records of functions from these concrete types to these concrete types paired
with new signal functions. The evaluation system for signal functions then
maintains this record for the signal function it is evaluating, calls the
correct function for the current input, actuates the outputs (using supplied
functions from the outputs to monadic actions) and replaces the record with the
newly supplied signal function record.

In Section~\ref{section:System_Design_and_Interface} we presented signal vectors
as a set of types. \footnote{In order to be completely principled, we should
isolate these types into their own {\em kind} (a sort of type of types);
however, the Haskell extension for this was far from stable at the time this
system was created.} Using the Glasgow Haskell Compiler's extension for
Generalized Algebraic Datatypes~\cite{Cheney2003,Xi2003,PeytonJones2006}, we
specify three concrete datatypes which are parameterized over signal vectors,
and represent information about a temporal point during evaluation.

@@

The next representation contains a constructor for {\tt SVEvent} nodes, carrying
a value typed by the type parameter of {\tt SVEvent}, and two constructors, each
of which carries a representation for one child of a {\tt SVAppend} node (the
left or right), and is parametric in the other. Thus, an instance of this
representation is a path from the root of a signal vector to an event node,
carrying exactly one value for the event. For instance, for an input occurrence
to the signal function:

\begin{code}
example :: (SVEvent Int :^: (SVEvent String :^: SVSignal Float))
\end{code}

might be:

\begin{code}
SVEventRight (SVEventLeft (SVEventOcc "Example"))
\end{code}

### 2012-11-30 Docs/IFL2012/introduction.tex
ifl draft edits

Almost all implementations of signal-function FRP to date~\cite{Courtney2001-1,Nilsson2002,Nilsson2005,Sculthorpe2011}
have used pull-based evaluation for both signals and events. This is due to the
ease of implementation of pull-based evaluation, and because the types used for
signal functions do not permit distinguishing signals and events, or
constructing only part of the input (for instance, an occurrence on one of the
event inputs). One signal function system~\cite{Jeffery2012} does permit the
user to push {\em segments} of signals, but, like the others, does not
distinguish events in a way that permits them to be separately pushed.

@@

Section~\ref{section:System_Design_and_Interface} describes design choices for the system,
and provides an overview of the interface. Section~\ref{section:Implementation}
describes how the system is implemented, and how the separation of evaluation
between events and signals is achieved. Section~\ref{section:Discussion} is a
discussion of the \textred{usefulness} \rn{more specific?} of our implementation. 
Section~\ref{section:Ongoing_and_Further_Work} describes the current and future
work on this system. Section~\ref{section:Related_Work} gives an overview of
related efforts. Section~\ref{section:Conclusion} concludes.

>>>>

Almost all implementations of signal-function FRP to date~\cite{Courtney2001-1,Nilsson2002,Nilsson2005,Sculthorpe2011}
have used pull-based evaluation for both signals and events. This is due to the
ease of implementation of pull-based evaluation, and because the types used for
signal functions do not permit distinguishing signals and events, or
constructing only part of the input (for instance, an occurrence on one of the
event inputs). One signal function system~\cite{Jeffrey2012} does permit the
user to push {\em segments} of signals, but, like the others, does not
distinguish events in a way that permits them to be separately pushed.

@@

Section~\ref{section:System_Design_and_Interface} describes design choices for the system,
and provides an overview of the interface. Section~\ref{section:Implementation}
describes how the system is implemented, and how the separation of evaluation
between events and signals is achieved. Section~\ref{section:Discussion} is a
discussion of the performance of our implementation in the implementation of a
network application. Section~\ref{section:Related_Work} gives an overview of
related efforts. Section~\ref{section:Conclusion} concludes.
